---
title: "Timeseries of Microcystis Lysed by a Virus Over a 36-hour Period"
author: "Eva Belevska"
format: gfm
editor: visual
---

## Time series of Microcystis Lysed by a Virus Over a 36-hour Period

To begin this data analysis, the data was loaded with the following code:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)

#The following code reads the csv file and loads the data frame. 
cell_density <- read_csv("data/cell_density.csv")
view(cell_density)
```

Many researchers have long asked why certain algal blooms happen at certain places. The plot generated below shows the cyanobacteria *Microcystis's* average optical density after they were lysed by a virus over a period of 36 hours. The amount of *Microcystis* that accumulates isn't only based on time and the amount of phosphate and nitrogen available - it is also dependent on the presence of removal mechanisms (@Wilhelm2020). In this case the removal mechanism is the virus, which is lysing and killing the bacterial cells. The graph shows that the optical density of the cyanobacteria decreases with the time that passes- as the virus is killing and lysing the bacterial cells. The inflection points for this data were generated with the following:

```{r, messages = FALSE, warning = FALSE}
#The following code finds the derivatives of the data.
CD_derivative <- cell_density |>
  mutate(derivative = c(NA, diff(Mean) / diff(Time)))
#The previously calculated derivatives are then used to find the inflection points.
inflection_points <- CD_derivative |>
  filter(derivative == 0 | (derivative * lag(derivative)) < 0)
view(inflection_points)

glimpse(inflection_points)
```

The inflection points show that the OD of the bacteria did not have an inflection point when there was a 0 rate of infection. This is because in this case there were no viruses in that sample to kill the cyanobacteria; therefore, it kept growing at a steady rate. However, for the other samples there seemed to be multiple inflection points. These points signify the times where the rate of optical density changed due to the virus lysing the cells. This shows that the more lysing that occurs, the less chance there is for a biofilm to form (@Wilhelm2020). The bacteria was also tested with differing rates of infection. This showed that the higher the rate of infection, the smaller the average optical density was.

The following code was used to generate a graph with the data:

```{r}
#A plot was created using the data frame that was previously loaded.
cell_density_graph_3 <- ggplot(data = cell_density,
  mapping = aes(x = Time, y = Mean, color = factor(VLPMOI)))+
#Points were added and a line was used to connect the points on the graph.
  geom_point()+
  geom_line()+
#The following adds labels to the graph. 
  labs(
    x = "Average Optical Density", 
    y = "Time (hours)", 
    title = "Average Microcystis Optical Density Over Time at 
                Different Multiplicities of Infection", 
    color = "Multiplicity of 
    Infection"
  )+
#Error bars were added to the graph. 
  geom_pointrange(aes(ymin= Mean - `Cell Density STD`, ymax = Mean + `Cell Density STD`), size = 0.1, position = position_dodge(width = 0))+
#A color brewer palette was used to accomodate colorblind people. 
  scale_color_brewer(palette = "Dark2")
  
print(cell_density_graph_3)
```

A graph of the different infection rates was generated as well, clearly showing the difference between each one using loess to smooth the data:

```{r}
#A plot was created using the data frame that was previously loaded. 
cell_density_graph <- ggplot(data = cell_density,
  mapping = aes(x = Time, y = Mean, color = factor(VLPMOI))) +
#Points were added and a line was used to connect the points on the graph. 
  geom_line()+
  geom_point()+
#The following adds labels to the graph. 
    labs(
    x = "Time (hours)", 
    y = "Average Optical Density", 
    title = "Average Microcystis Optical Density Over Time at 
                Different Multiplicities of Infection", 
    color = "Multiplicity of 
    Infection")+
#This creates a graph for each multiplicity of infection.
  facet_wrap(~factor(VLPMOI))+
#This smooths the data. 
  geom_smooth(method = loess)+
#It is important to use a color scheme that takes colorblind people into account, so the a color brewer palette was utilized. 
  scale_color_brewer(palette = "Dark2")

print(cell_density_graph)
```

## References:

Wilhelm, S. W., Bullerjahn, G. S., & McKay, M. L. (2020). The Complicated and Confusing Ecology of Microcystis Blooms. https://www.bgsu.edu/content/dam/BGSU/great-lakes-center/Wilhelm-2020.pdf
