---
title: "Bad Graph"
author: "Eva Belevska"
format: gfm
editor: visual
---

# Homework 7 

## Bad Plot

```{r, message=FALSE, warning = FALSE}
library(tidyverse)
library(palmerpenguins)

ggplot(data = penguins) +
  geom_point(
    mapping = aes(x = bill_length_mm, y = bill_depth_mm, size = island, alpha = island), data = penguins, 
    color = "yellow"
  )
```

This is a bad plot because the color chosen for the points is hard to look at. According to Wilkes a good plot will not have "junk". This plot is somewhat filled with "junk" because the size and opaqueness of the dots both represent the island type. Additionally,the opacity of the dots also makes it hard to read because it is hard to tell which points are which shade of yellow because they all seem to clump together and overlap. There is no way to tell where one dot ends and another begins. Furthermore, Wilkes states that how bright an object looks depends on how bright its surroundings are. this could also be applied to this graph because if a really bright dot is which, represents the Torgersen island, is close to the dot representing Dream island, it could make the duller dot seem less opaque and it could be mistaken for Biscoe island. 
The intent of this plot was to compare the bill length against the bill depth on penguins, and to also compare how this varies between the three islands. However, it fails to do present this data in a way that is easy to read. 

## Fixed Plot

```{r, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x=bill_length_mm, y=bill_depth_mm, color=island)) +
geom_point(size=1.5) +
scale_color_brewer(palette = "Dark2")
```

This plot is much easier to read than the previous plot. The islands are only represented by one thing (different colors) rather than two (opacity and size). Furthermore, the colors that were used are not as bright as the yellow in the previous graph. Additionally, Healy mentions that colorblind people may not be able to read some graphs, so the color brewer palette "Dark2" was used to accommodate for this. 