---
title: "HMK 5: Reading and tidying data"
format: gfm
---

# Reading

-   [R4DS Chapters 6-9](https://r4ds.hadley.nz/data-import)

# Data import

## Q1:

-   Create a directory, within your main class directory, called `data`.
    -   Note: in general, you should store raw data in a directory called `data`.
-   Download the example file for Ch 9 here. Save it inside the `ddata` directory.
-   Use `read_csv()` to read the file to an R data frame. Follow the instructions in Ch 9 to format it properly. Follow the directions in Ch 9 to make sure the following are true:
    -   Column names should be *syntactic*, meaning they don't contain spaces.
    -   N/A values should be represented with the R value `NA`, not the character "N/A".
    -   Data types (character vs factor vs numeric) should be appropriate.
```{r, message=FALSE}
library(tidyverse)
students <- read_csv("Data/students.csv", na = c("N/A", ""))
students |>
  rename(
    student_id = `Student ID`,
    full_name = `Full Name`, 
    favourite_food = `favourite.food`, 
    meal_plan = `mealPlan`,
    age = `AGE`
  ) |>
  mutate(
    meal_plan = factor(meal_plan),
    age = parse_number(if_else(age == "five", "5", age))
    )
```


## Q2

Find (or make) a data file of your own, in text format. Read it into a well-formatted data frame.

```{r, message=FALSE}
lost_pets <- read_csv("Data/Lost__found__adoptable_pets.csv")
lost_pets |>
  rename(Impound_No = `impound_no`) |>
  mutate(Record_Type = factor(Record_Type))
```


# Tidying

Download the data set available at <https://tiny.utk.edu/MICR_575_hmk_05>, and save it to your `data` folder. **This is a real data set:** it is the output from the evaluation forms for student colloquium speakers in the Microbiology department. I have eliminated a few columns, changed some of the scores, and edited comments, to protect student privacy, but the output is real.

First, open this .csv file with Microsoft Excel or a text reading app, to get a sense of the structure of the document. It is weird.

Why is the file formatted so inconveniently? I have no idea, but I do know that this is about an average level of inconvenient formatting for real data sets you will find in the wild.

*Note: In theory, you can pass a URL to `read_csv()` and read the file directly from the internet. In practice, that doesn't seem to work for this file. So you'll want to download this file to your hard drive.*

## Q3a

Next, use `read_csv()` to read the data into a data frame. Note that you'll need to make use of some of the optional arguments. Use `?read_csv` to see what they are.

*If you are struggling with this task, email me for hints.*

As we discussed in class, the correct shape depends on what you want to do with the data. Use `pivot_longer()` to make the data frame longer, in a way that makes sense.

```{r, message = FALSE}
headers <- names(read_csv("Data/colloquium_assessment.csv", n_max = 0))
                 
colloquium <- read_csv("Data/colloquium_assessment.csv", col_names = headers,skip = 6)

colloquium_clean <- colloquium |>
  filter(!apply(colloquium == "NA", 1, all)) |>
  select(-contains("RecipientLastName")) |>
  select(-contains("RecipientFirstName")) |>
  select(-contains("RecipientEmail")) |>
  rename("Comments" = "Q12") |>
  rename("Student's Track" = "Q4") |>
  rename("Year" = "Q5") |>
  rename("Are_You" = "Q6") |> 
  rename("Presentation_Score" = "Q11")|>
  pivot_longer(
    cols = starts_with("Q"), 
    names_to = "questions", 
    values_to = "score",
    values_drop_na = TRUE
  )
```


## Q3b

Finally, calculate this student's average score for each of questions 7-10.

```{r, message = FALSE}
average_score <- colloquium_clean |>
  group_by(questions) |>
  summarise(mean = mean(score)) |>
  filter(questions == "Q7"|questions == "Q8"|questions == "Q9"|questions == "Q10")
```


## Important note about file paths in Quarto documents

When you render a Quarto document, RStudio spins up a new instance of R, which is separate from the instance of R that you cna interact with. The working directory for this instance of R is whatever directory your Quarto document is saved in.

If your quarto document is saved in the same directory as your RStudio project (e.g., `MICR_475`), then there is no difference between your interactive working directory and the working directory for your Quarto document.

However, if your homeworks are saved in a `HMK` directory, then the Quarto working directory will be `HMK`. To access the saved `.csv` file, `read_csv()` will need to look *up* one directory and then go back *down* into `HMK`. `..` means "up one directory", so you would need to use `read_csv("../colloquium_assessment.csv")` instead of `read_csv("colloquium_assessment.csv")`.
